---
title: '"Data Transformation With dplyr"'
author: '2206 :Divya Chaudhari , 2225: mayuri patil'
date: "10/12/2022"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why do I use dplyr?
* Great for data exploration and transformation
* Intuitive to write and easy to read, especially when using the “chaining” syntax (covered below)
* Fast on data frames

## dplyr functionality
* Five basic verbs: filter, select, arrange, mutate, summarise (plus group_by)
* Can work with data stored in databases and data tables
* Joins: inner join, left join, semi-join, full-join 

--------


## attched library

```{r}
library(dplyr)
library(hflights)
data(hflights)
head(hflights)
flights <- tbl_df(hflights)
flights

```
-----------

## filter: Keep rows matching criteria
* Base R approach to filtering forces you to repeat the data frame’s name
* dplyr approach is simpler to write and read
* Command structure (for all dplyr verbs):
– first argument is a data frame
– return value is a data frame
– nothing is modified in place
* Note: dplyr generally does not preserve row names

```{r echo=TRUE}
filter(flights, Month==1, DayofMonth==1)
# use pipe for OR condition
filter(flights, UniqueCarrier=="AA" | UniqueCarrier=="UA")
# you can also use %in% operator
filter(flights, UniqueCarrier %in% c("AA", "UA"))
```
-----

## select: Pick columns by name
* Base R approach is awkward to type and to read
* dplyr approach uses similar syntax to filter
* Like a SELECT in SQL

```{r}
select(flights, DepTime, ArrTime, FlightNum)
select(flights, Year:DayofMonth, contains("Taxi"), contains("Delay"))
```
---

# “Chaining” or “Pipelining”
* Usual way to perform multiple operations in one line is by nesting
* Can write commands in a natural order by using the %>% infix operator (which can be pronounced as
“then”)

```{r}
flights %>%
select(UniqueCarrier, DepDelay) %>%
filter(DepDelay > 60)
```

-----

# arrange: Reorder rows
*which is used to sort rows in a dataframe based on attributes held by particular columns

```{r}
flights %>%
  arrange(UniqueCarrier, DepDelay)
```
-----

## mutate: Add new variables
*  which is used to sort rows in a dataframe based on attributes held by particular columns

```{r}
flights %>%
select(Distance, AirTime) %>%
mutate(Speed = Distance/AirTime*60)
```
-------

### summarise: Reduce variables to values
* which is used to collapse values from a dataframe into a single summary.

```{r}
flights %>%
group_by(Dest) %>%
summarise(avg_delay = mean(ArrDelay, na.rm=TRUE))
```

------

### The four previous join functions (i.e. inner_join, left_join, right_join, and full_join) are so called mutating joins. Mutating joins combine variables from the two data sources.

```{r}
data1 <- data.frame(ID = 1:2,                      # Create first example data frame
                    X1 = c("a1", "a2"),
                    stringsAsFactors = FALSE)
data2 <- data.frame(ID = 2:3,                      # Create second example data frame
                    X2 = c("b1", "b2"),
                    stringsAsFactors = FALSE)
```

## inner_join dplyr R Function
* In order to merge our data based on inner_join, we simply have to specify the names of our two data frames (i.e. data1 and data2) and the column based on which we want to merge (i.e. the column ID)
```{r}
inner_join(data1, data2, by = "ID")  

```

## left_join dplyr R Function
```{r}
left_join(data1, data2, by = "ID") # Apply left_join dplyr function
```
## right_join dplyr R Function
```{r}
right_join(data1, data2, by = "ID") 
```
## semi_join dplyr R Function
```{r}
semi_join(data1, data2, by = "ID")                 # Apply semi_join dplyr function
```
## full_join dplyr R Function
```{r}
full_join(data1, data2, by = "ID")                 # Apply full_join dplyr function
```


