---
title: <span style="color:red">"Apply function with purrr"

author: "Tejas 2230, Darshan 2204, Bhushan 2218"

date: "14 DEC 2022"

output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


---

#<span style="color:red">**Introduction**

In this topic we talked about how important it is to reduce duplication
in your code by creating functions instead of copying and
pasting. Reducing code duplication has three main benefits:

• It’s easier to see the intent of your code, because your eyes are
drawn to what’s different, not what stays the same.

• It’s easier to respond to changes in requirements. As your needs
change, you only need to make changes in one place, rather
than remembering to change every place that you copied and
pasted the code.

• You’re likely to have fewer bugs because each line of code is used
in more places


---

Every for loop has three components:

output output <- vector("double", length(x))
Before you start the loop, you must always allocate sufficient
space for the output. This is very important for efficiency: if you
grow the for loop at each iteration using c() (for example), your
for loop will be very slow.

A general way of creating an empty vector of given length is the
vector() function. It has two arguments: the type of the vector
(“logical,” “integer,” “double,” “character,” etc.) and the length of
the vector.

sequence i in seq_along(df)

This determines what to loop over: each run of the for loop will
assign i to a different value from seq_along(df). It’s useful to
think of i as a pronoun, like “it.”

```{r,echo=TRUE}
library(tidyverse)
df=tibble(a=rnorm(10),b=rnorm(10),c=rnorm(10),d=rnorm(10))
df
med=median(df$a)
med


median(df$a)
median(df$b)
median(df$c)
median(df$d)


output=vector("double",ncol(df))
for (i in seq_along(df)) {
  output[[i]]=median(df[[i]])
}
output
```

---

#**The Map Functions**

The pattern of looping over a vector, doing something to each element,
and saving the results is so common that the purrr package
provides a family of functions to do it for you. There is one function
for each type of output:

• map() makes a list.

• map_lgl() makes a logical vector.

• map_int() makes an integer vector.

• map_dbl() makes a double vector.

• map_chr() makes a character vector.


   Each function takes a vector as input, applies a function to each
piece, and then returns a new vector that’s the same length (and has
the same names) as the input. The type of the vector is determined
by the suffix to the map function.

```{r,echo=TRUE}
map_dbl(df, median)
map_dbl(df, mean)
```
```{r , echo=TRUE}
library("tidyverse")
p=c(1,2,3,4,5,6,7)
q=c(8,9,10,3,5,6,7)
r=c(3,9,6,4,1,8,5)
pmap(list(p,q,r),sum)

```

---

Compared to using a for loop, focus is on the operation being performed
(i.e., mean(), median(), sd()), not the bookkeeping required
to loop over every element and store the output. This is even more
apparent if we use the pipe:

```{r,echo=TRUE}
df %>% map_dbl(mean)
df %>% map_dbl(median)
df %>% map_dbl(sd)
```


---


#**Linear model**

There are a few shortcuts that you can use with .f in order to save a
little typing. Imagine you want to fit a linear model to each group in
a dataset. The following toy example splits up the mtcars dataset
into three pieces (one for each value of cylinder) and fits the same
linear model to each piece:
```{r,echo=TRUE}
library(dplyr)
df5=mtcars
mtcars%>%map_dbl(sum)
mtcars%>%
  split(mtcars$hp)%>%
  map(~lm(mpg~wt,data = df5))%>%
  #map_dfr(~ as.data.frame(t(as.matrix(coef(.)))))
  map(summary)
  
model=mtcars%>% split(mtcars$cyl)%>%map(function(df5) lm(mpg~wt,data=df5))
model
```


---


#**Base R**

If you’re familiar with the apply family of functions in base R, you
might have noticed some similarities with the purrr functions:

• lapply() is basically identical to map(), except that map() is
consistent with all the other functions in purrr, and you can use
the shortcuts for .f.

• Base sapply() is a wrapper around lapply() that automatically
simplifies the output. This is useful for interactive work but is
problematic in a function because you never know what sort of
output you’ll get:

```{r,echo=TRUE}
x1 = list(
  c(0.27, 0.37, 0.57, 0.91, 0.20),
  c(0.90, 0.94, 0.66, 0.63, 0.06),
  c(0.21, 0.18, 0.69, 0.38, 0.77)
)
x2 =list(
  c(0.50, 0.72, 0.99, 0.38, 0.78),
  c(0.93, 0.21, 0.65, 0.13, 0.27),
  c(0.39, 0.01, 0.38, 0.87, 0.34))
threshould=function(x,cutoff=0.8) x[x>cutoff]
x1%>%sapply(threshould)%>%str()
x2%>%sapply(threshould)%>%str()

```


---

#**THANK YOU**
  
