---
title: "NON-PARAMETRIC TESTS USING R MARKDOWN"
author: "2201-Prachi Badgujar, 2216- Sayali Nikhumbh ,2219-Damini Patil"
date: "14/12/2022"
output: slidy_presentation
---
# What is non-parametric test?

Data do not follow any specific distribution and no assumptions about the population are Made

Data measured on any scale

# Parameter

Parameter is any numerical quntity that characterizes a given population or some aspects of it. Most common statistics parameters are mean, median, mode, standard deviation.

# Commonaly used tests

-   Chi-Square tests

-   Sign test

-   Wilcoxon Signed-ranks test

-   Mann-Whitney U or Wilcoxon rank sum test

-   Shapiro wilk tests

-   Kolmogorav-Smirov (K-S) test

------------------------------------------------------------------------

# Chi-Square

-   First used by Karl Pearsons

-   Simplest and Most Widely used Non-Parametric test in statistical work

-   It is used for testing relationship between categorical variable

-   compare the observed values in your data to the expected values that you would see if the null hypothesis is true

-   calculated using :

     $\chi^2$= $\sum((O-E)^2/E)$

    where,  O=observed frequencies
    E=expected frequency


# Application of chi-square test

-   Test of association(Smoking & Cancer, treatment & outcome of diesease , vaccination & immunity).

-   Test of proportion (compare frequencies of diabetics & non-diabetics in groups).

-   The chi-square for goodness of fit (determined if actual numbers are similar to the expected\\theoretical numbers).

---


# Example
```{r}
x=trunc(5 * runif(100));x
table(x)
chisq.test(table(x))
fit=lm(weight~height,data=women);fit
summary(fit)
residuals(fit)
```

---


```{r}
plot(women$weight~women$height)
abline(fit,col="blue")

```

---

# Sign test

- used for paired data, can be ordinal or continuous.

- simple and easy to interpret. based sign(plus or minus)

- Makes no assumptions about distribution of the .

- To evaluate H0 we only need to know the signs of the differences .

- if the half differences are positive and half are negative then median=0(H0 is true).

- If the signs are unbalanced ,then that is evidences against H0

---


## sign test for one sample:
    Qn: A manufacturering plant claim to produce widgets that weigh 50 pounds .to test this, an inspector goes out to the plant and randomly measures the weigh
    of 20 widgets. determine if the median weights is significantly different from 50 pounds.
```{r}
x=c(51,52,52,50,49,47,52,55,48,49,50,51,51,51,49,46,51,52,47,52);x
md=50  ### median given
sp=length(x[x<md]);sp
sn=length(x[x>md]);sn
n=sp+sn;n
pv=binom.test(sp,n,0.5);pv

```

 ## Conclusion : Here sp is the value of s+ and pv is the p- value. if given level of significance is less than p-value we accept the H0 here alpha=0.05. Hence accept H0 so population median of x is 50.
 
 
 ---
 
# Sign test for two samples
  Qn : An I.Q test was administered 5 students before and after they were class. The results are given below
```{r}
students=c(1,2,3,4,5)
before =c(110,120,123,132,125)
after=c(120,118,125,136,121)
data.frame(students,before,after)
d=before-after
sp=length(d[d>0]);sp
sn=length(d[d<0]);sn
n=sp+sn;n
pv=pbinom(sp,n,0.5);pv
```
## Conclusion: 
 Here sp is value of s+ and pv is the p-value. If given level of significance is less than p-value,
we accept Ho Here alpha=0.05. hence accept Ho. So there is no increase in I.Q. after traning program. 

---

 
# Wilcoxon Singed Rank Test

- Wilcoxon singed rank test is similar to sign test but sign test consider only direction & wilcoxon signed rank test consider direction as well as magnitude.

- This test is used to determine whether two dependent samples were selected from population having the same distribution.

- it also known as t-test for matched pairs or t-test for dependent samples.

- There are two types of wilcoxon singed rank test.


---

# One-sample wilcoxon signed rank test

  The given data set containing the weight of 10 rabbits. If the median weight of the rabbits differs from 25g?
  
```{r}
 set.seed(1234)
myData=data.frame(name=paste0(rep("R",10),1:10),weight=round(rnorm(10,30,2),1))
print(myData)
wilcox.test(myData$weight,mu=25)
```

  # Conclusion: In the given output , p-value of the test is 0.005793, which is less than significance level alpha=0.05.So we can reject null hypothesis.There for we conclude that the average weight of the rabbit is significantly different from 25g. 



# Example for Two-sample wilcoxon signed rank test

  The following are the weights in Kg of 16 persons.Before and After a certain of four weeks are given. Test whether the weight reducing diet is effective at 0.01 level of significance.  
  
```{r}
 before=c(80.7,78.1,82.4,86.8,90.1,75.9,74.5,72.2,74.3,87.5,86.1,85.75,81.89,88.3,83.33,73)
 after=c(75.1,70.3,74.4,80.7,75.9,60.12,62.15,75,69.8,89,70.2,76.4,72.7,76.9,79.11,66.16)
 d=before-after
 wilcox.test(d,mu=0,alternative = "less")
```

 # Conclusion: In the given output, p-value is greater than given significance level of alpha. So we accept null hypothesis. There for we conclude that weight reducing diet is effective 
 
 ---
 
## Mann- Whitney U test
 - This test is similar to wilcoxon signed rank test expect that the sample are independent and not paired.
 
 - In this test the null hypothesis of the population means are same for two groups.
 
 - Rank the combined data value for two groups. Then find average rank of each groups.
  
- Then U value is calculated using the formula:
     $\ U=N1*N2+(Nx(Nx+1)/2)-Rx$
     
## Example for Mann-Whitney tests
qn: The following are the lengths of time (in minute) spend in the operating room by 20 patients undergoing the some operative procedure.
```{r}
A=c(35,30,33,39,41,29,30,36,45,40,31);A
B=c(45,38,42,50,48,51,32,37,46);B
wilcox.test(A,B,alter="greater")
```
## conclusion:
  In the given output, p-value is greater than given significance level of alpha. So we accept null hypothesis. Therefor so the patients in the hospital B tend to be in the operating room longer than patients in hospital A for a some operative procedure.
  
---

## Shapiro wilk tests

- It examines how close the sample data fit to a normal distributions.

- If the p-value of the test if less than 0.05, then we reject the null hypothesis at 5% significance and conclude that your data is non-normal.

- The shapiro wilk test is more appropriate for the small size samples(<50).

---


## Examples for Shapiro wilk tests
```{r}
library("PASWR")
 attach(Fertilize)
 View(Fertilize)
shapiro.test(self)
plot(density(cross,self))

n=length(cross)
m=length(self)

g=as.factor(c(rep("cross",n),rep("self",m)))
k=c(cross,self)


```


## conclusion:
There is significant difference in both the variables.

---

## Kolmogorav-Smirnov test
- The kolmogorav-smirnov test is used for goodness of fit
- Is used to decide if a sample comes from a population with specific distribution.
- The formula for K-S test can be given as
$\ Dn=supx | Fn(x)-F(x)|$
```{r}
library("dgof")
x=rnorm(50)
y=runif(30)
ks.test(x,y)
plot(ecdf(x),
     xlim=range(c(x,y)),
     col="blue")
plot(ecdf(y),
    add=TRUE,
    lty= "dashed",
    col= "red")

```
## conclusion:
  Being quite sensitive to the difference of ahape and location of the emipirical cumulative distribution of the chosen two samples,The two sample k-s test is efficient and one of the most general and usefull non parameteric test. hence from the we can see the difference between two samples.

---



 

## Advantages of non- parametric tests
- These tests are distribution free.

- Easier to calculated and less time consuming than parametric tests when sample size is small.

- Can be used with any type of data.

- Many non-parametric methods make it possible to work with very small samples, particularly helpful in collecting pilot study data or medical researcher working with a rare        disease.

---

## Limitations of non-parametric methods

- Statistical methods which requires no assumptions about populations are usually less efficient.

- As the sample size get larger, data manipulatios required for non-parametric tests becomes tests becomes laborious.

- A collection of tabulated critical values for a variety of non-parametric tests under situations dealing with various sample sizes is not readily available.


 
 
 
 
 


